The tree line ecotone punctuates the abundance and β-diversity of plants, pollinators, and their interactions

### 2021-09-02 code/ms1_abundance_v2.Rmd
beginning ms1 revision

# 1. Prep work
## 1.1 Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(patchwork)
library(tidymv)
library(vegan)
library(tidyverse)
library(lubridate)
```

@@

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male",
         !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

@@

## 1.4 Calculate FL and BB abundance
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them. 
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)
```

## 1.5 Calculate binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:14, values_to = "visited", names_to = "bb.sp")

weighted_fl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))
```

@@

>>>>

# 1. Prepare environment
## 1.1 Load packages
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(patchwork)
library(tidymv)
library(vegan)
library(tidyverse)
library(lubridate)
```

@@

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
# net <- read_csv("../data/processed_data/network.csv") %>%
#   filter(caste != "male",
#          !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
#   dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% 
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

@@

# 2. Process data
## 2.1 Calculate transect-wise FL and BB abundance
```{r echo=FALSE, include=TRUE, message=FALSE}
# Survey transects
survey_transects_site <- survey %>%
  group_by(site) %>%
  summarize(survey.transects = length(unique(date)))

# BB transects
net_transects_site <- net %>%
  group_by(site) %>%
  summarize(net.transects = length(unique(date)))

# Per-transect BB abundance
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them. 
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)

### Bumble bee richness per site
bb_rich_site <- net %>%
  group_by(site) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Peak and mean BB abundance per site
bb_abund_site <- bb_abund_sp %>%
  group_by(site, date) %>%
  summarize(bb.total.abund = sum(bb.abund)) %>%
  group_by(site) %>%
  summarize(bb.peakabund = max(bb.total.abund)) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Floral richness per site
fl_rich_site <- survey %>%
  group_by(site) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Peak floral abundance per site
fl_abund_site <- survey %>%
  group_by(site, date) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  group_by(site) %>%
  summarize(fl.peakabund = max(fl.abund),
            fl.meanabund = mean(fl.abund),
            fl.medabund = median(fl.abund)) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Interaction richness per site
interaction_rich_site <- net %>%
  group_by(site) %>%
  unite(interaction, c(bb.sp, plant.sp), sep = "_") %>%
  summarize(interaction.rich = length(unique(interaction))) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

rich_abund_collated <- bb_rich_site %>% 
  full_join(fl_rich_site) %>%
  full_join(bb_abund_site) %>% 
  full_join(fl_abund_site) %>%
  full_join(interaction_rich_site)

rich_abund_collated.long <- rich_abund_collated %>%
  pivot_longer(cols = c(bb.rich, fl.rich, bb.peakabund, 
                        fl.peakabund, fl.meanabund, 
                        fl.medabund, interaction.rich)) %>%
  separate(name, c("group", "metric"), sep = "\\.") %>%
  group_by(group, metric) %>%
  mutate(z = scale(value))

ggplot(rich_abund_collated.long, aes(elev.mean, z, color = group)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~metric) +
  theme_light()

# Exploring some bivariate relations

ggplot(rich_abund_collated, aes(fl.rich, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_collated, aes(fl.peakabund, fl.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_collated, aes(fl.peakabund, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_collated, aes(fl.peakabund, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm") 
```

## 2.2 Summarize

@@

## 2.3 Mean total floral cover (not broken down by bumble bee species) ~ elevation * time
```{r}
fl_abund_total <- fl_abund_sp %>%
  group_by(site, date) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

bb_abund_total <- net %>%
  group_by(site, date) %>%
  summarize(bb.abund = n()) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

fl_rich_total <- fl_abund_sp %>%
  filter(fl.abund > 0) %>%
  group_by(site, date) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

bb_rich_total <- net %>%
  group_by(site, date) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

fl_rich_total <- fl_abund_sp %>%
  filter(fl.abund > 0) %>%
  group_by(site, date) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

ggplot(fl_abund_total, aes(fl.abund)) +
  geom_histogram() +
  facet_wrap(~site)

ggplot(fl_rich_total, aes(fl.rich)) +
  geom_histogram() +
  facet_wrap(~site)

### Abundance
total_fl_abund_GAM <- gam(log(fl.abund + 0.01) ~ 
                             year +
                             management +
                             s(site, bs = "re") +
                             te(yday, elev.mean, by = year, bs=c("gp","tp")),
                           family = "gaussian",
                           method = "REML",
                           data = fl_abund_total) %>% getViz()

check.gamViz(total_fl_abund_GAM)
summary(total_fl_abund_GAM)
plot.gamViz(total_fl_abund_GAM, allTerms = TRUE)

total_bb_abund_GAM <- gam(bb.abund ~ 
                             year +
                             management +
                             s(site, bs = "re") +
                             te(yday, elev.mean, by = year, bs=c("gp","tp")),
                           family = "nb",
                           method = "REML",
                           data = bb_abund_total) %>% getViz()

check.gamViz(total_bb_abund_GAM)
summary(total_bb_abund_GAM)
plot.gamViz(total_bb_abund_GAM, allTerms = TRUE)

### Richness

total_fl_rich_GAM <- gam(fl.rich ~ 
                            year +
                            management +
                            s(site, bs = "re") +
                            te(yday, elev.mean, by = year, bs=c("gp","tp")),
                          family = "poisson",
                          method = "REML",
                          data = fl_rich_total) %>% getViz()

check.gamViz(total_fl_rich_GAM)
summary(total_fl_rich_GAM)
plot.gamViz(total_fl_rich_GAM, allTerms = TRUE)

total_bb_rich_GAM <- gam(bb.rich ~ 
                            year +
                            management +
                            s(site, bs = "re") +
                            te(yday, elev.mean, by = year, bs=c("gp","tp")),
                          family = "nb",
                          method = "REML",
                          data = bb_rich_total) %>% getViz()

check.gamViz(total_bb_rich_GAM)
summary(total_bb_rich_GAM)
plot.gamViz(total_bb_rich_GAM, allTerms = TRUE)
```

### 2021-09-02 code/ms1_beta_diversity.Rmd
beginning ms1 revision

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% #
  mutate(site = factor(site),
         tree_line = factor(tree_line)) %>% 
  semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)

# Write to csv for archiving
write_csv(site_data, "../Dryad_ms1/site_data.csv")
```

@@

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

# Write to csv for archiving
write_csv(net, "../Dryad_ms1//visitation_data.csv")
write_csv(survey, "../Dryad_ms1//floral_survey_data.csv")
```

@@

# Write to csv for archiving
write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
```

@@

# BB + FL elevation diff
g1 <- ggplot(filter(species_beta, 
                    site1 != site2), # no within-site comparisons
             aes(elev.diff, beta, color = group)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = FALSE) +
  xlab("Elevation difference (m)") +
  ylab("β-diversity") +
  theme_light()

@@

# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
png("../output/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
(g1 | (g2 / g3)) + 
  plot_annotation(tag_levels = c("A", "B", "C")) + 
  theme_light() 
dev.off()

pdf("../output/beta_stack.pdf", width = 6.5, height = 4.5)
(g1 | (g2 / g3)) + 
  plot_annotation(tag_levels = c("A", "B", "C")) + 
  theme_light() 
dev.off()
```

@@

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Prep sitepairs
sitepair_WN <- interaction_gdm_prep(net)

# Run GDM analysis
gdm_WN <- gdm(sitepair_WN, geo = FALSE)
summary(gdm_WN)
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)

# Plot
plot(gdm_WN)

@@

# Call GDMs
gdm_survey <- gdm(sitepair_survey, geo = FALSE)
summary(gdm_survey)
gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE)

# Plot
plot(gdm_survey)

@@

# Call GDMs
gdm_BB <- gdm(sitepair_BB, geo = FALSE)
summary(gdm_BB)
gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = FALSE)

# Plot
plot(gdm_BB)

@@

### Plot GDM splines
```{r}
# Interaction model
gdm_WN_splines1 <- isplineExtract(gdm_WN)$x %>%
  as.data.frame() %>%
  select(elevation = elev.mean)

gdm_WN_splines2 <- isplineExtract(gdm_WN)$y %>%
  as.data.frame() %>%
  select(response = elev.mean)

@@

# Flora model
gdm_survey_splines1 <- isplineExtract(gdm_survey)$x %>%
  as.data.frame() %>%
  select(elevation = elev.mean)

gdm_survey_splines2 <- isplineExtract(gdm_survey)$y %>%
  as.data.frame() %>%
  select(response = elev.mean)

@@

# BB model
gdm_BB_splines1 <- isplineExtract(gdm_BB)$x %>%
  as.data.frame() %>%
  select(elevation = elev.mean)

gdm_BB_splines2 <- isplineExtract(gdm_BB)$y %>%
  as.data.frame() %>%
  select(response = elev.mean)

@@

tree_line <- site_data %>%
  select(elevation = elev.mean, tree_line)

@@

ggsave("../output/gdm_splines.pdf")
ggsave("../output/gdm_splines.png")

>>>>

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% #
  mutate(site = factor(site),
         tree_line = factor(tree_line)) %>% 
  semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)
```

@@

survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant spp. (anytime, anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) 
```

@@

# Write to csv for archiving
#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
```

@@

# BB + FL elevation diff
g1 <- ggplot(filter(species_beta, 
                    site1 != site2), # no within-site comparisons
             aes(elev.diff, beta, color = group)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"), 
              se = FALSE) +
  xlab("Elevation difference (m)") +
  ylab("β-diversity (proportional species turnover)") +
  theme_light()

@@

# Patchwork plot for comparing species beta-diversity with interaction beta-diversity
png("../ms_1/figures/beta_stack.png", width = 6.5, height = 4.5, units = "in", res = 300)
(g1 | (g2 / g3)) + 
  plot_annotation(tag_levels = c("A", "B", "C")) + 
  theme_light() 
dev.off()

pdf("../ms_1/figures/beta_stack.pdf", width = 6.5, height = 4.5)
(g1 | (g2 / g3)) + 
  plot_annotation(tag_levels = c("A", "B", "C")) + 
  theme_light() 
dev.off()
```

@@

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Prep sitepairs
sitepair_WN <- interaction_gdm_prep(net)
sitepair_WNb <- interaction_gdm_prep2(net)

# Run GDM analysis
gdm_WN6 <- gdm(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm_WN6b <- gdm(sitepair_WNb, geo = FALSE, splines = c(6, 6, 6, 6, 6))

gdm_WN6 <- gdm(sitepair_WN, geo = TRUE, splines = c(6, 6, 6, 6, 6, 6))

summary(gdm_WN)

gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6))
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6))

# gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE, splines = c(6, 6, 6, 6))

# Plot
plot(gdm_WN6)
plot(gdm_WN6b)

@@

# Call GDMs
gdm_survey10 <- gdm(sitepair_survey, geo = TRUE, splines = c(10, 10, 10, 10, 10, 10))
gdm_survey6 <- gdm(sitepair_survey, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm_survey5 <- gdm(sitepair_survey, geo = TRUE, splines = c(5, 5, 5, 5, 5, 5))
gdm_survey4 <- gdm(sitepair_survey, geo = TRUE, splines = c(4, 4, 4, 4, 4, 4))

summary(gdm_survey)
gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE, splines = c(6, 6, 6, 6, 6, 6))

# Plot
plot(gdm_survey10)
plot(gdm_survey6)
plot(gdm_survey5)
plot(gdm_survey4)

@@

# Call GDMs
gdm_BB6 <- gdm(sitepair_BB, geo = FALSE, splines = c(6, 6, 6, 6, 6))
gdm_BB5 <- gdm(sitepair_BB, geo = FALSE, splines = c(5, 5, 5, 5, 5, 5))
gdm_BB4 <- gdm(sitepair_BB, geo = TRUE, splines = c(4, 4, 4, 4, 4, 4))

summary(gdm_BB)
gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = TRUE, splines = c(6, 6, 6, 6, 6, 6))

# Plot
plot(gdm_BB6)
plot(gdm_BB5)
plot(gdm_BB4)

@@

### Fixed isplineExtract()
```{r}
##Extracts Ispline data from a gdm model
isplineExtract2 <- function (model){
  ###########################
  #model = gdmOb
  ###########################
  ##error checking
  ##checks to make sure a gdm model is given
  if(!is(model, "gdm")){
    stop("model argument must be of class = 'gdm'.")
  }

  ##Collects or sets simple data
  options(warn.FPU = FALSE)
  PSAMPLE <- 200
  preddata <- rep(0, times = PSAMPLE)
  pn <- model$predictors
  nPreds <- length(pn)
  yDat <- xDat <- matrix(0,PSAMPLE,nPreds)
  colnames(yDat) <- colnames(xDat) <- pn
  pmin <- 1
  pmax <- PSAMPLE

  ##cycles through each prodictor and fills the spline matrices
  splineindex <- 1
  for (i in 1:nPreds){ 
    #i<-1
    numsplines <- model$splines[i]
    z <- .C("GetPredictorPlotData", 
            pdata = as.double(preddata), 
            as.integer(PSAMPLE), 
            as.double(model$coefficients[splineindex:(splineindex + numsplines - 1)]), 
            as.double(model$knots[splineindex:(splineindex + numsplines - 1)]), 
            as.integer(numsplines),
            PACKAGE = "gdm")
    yDat[,i] <- z$pdata
    pmin <- pmin + PSAMPLE
    pmax <- pmax + PSAMPLE
    xDat[,i] <-  seq(from=model$knots[[(i*numsplines)-(numsplines-1)]],to=model$knots[[(i*numsplines)]], length=PSAMPLE)
    splineindex <- splineindex + numsplines
  }

  ##lists and returns matrices
  outData <- list(x=xDat,y=yDat)
  return(outData)
}
```

### Plot GDM splines
```{r}
# Interaction model
gdm_WN_splines1 <- isplineExtract2(gdm_WN6)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

gdm_WN_splines2 <- isplineExtract2(gdm_WN6)$y %>%
  as.data.frame() %>%
  dplyr::select(response = elev.mean)

@@

# Flora model
gdm_survey_splines1 <- isplineExtract2(gdm_survey6)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

gdm_survey_splines2 <- isplineExtract2(gdm_survey6)$y %>%
  as.data.frame() %>%
  dplyr::select(response = elev.mean)

@@

# BB model
gdm_BB_splines1 <- isplineExtract2(gdm_BB6)$x %>%
  as.data.frame() %>%
  dplyr::select(elevation = elev.mean)

gdm_BB_splines2 <- isplineExtract2(gdm_BB6)$y %>%
  as.data.frame() %>%
  dplyr::select(response = elev.mean)

@@

tree_line <- site_data %>%
  dplyr::select(elevation = elev.mean, tree_line)

@@

ggsave("../output/gdm_splines2.pdf")
ggsave("../output/gdm_splines2.png")

### 2021-09-02 code/ms1_sampling.Rmd
beginning ms1 revision

---
title: "Appendix S1"
output: pdf_document
---

# Packages
```{r message=FALSE, warning=FALSE}
library(mgcViz)
library(lubridate)
library(tidyverse)
```

@@

## Visualize
```{r message=FALSE, warning=FALSE}
ggplot(visitation_sampling, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Observations") +
  facet_wrap(~year(date)) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per study year")

ggplot(bb_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Sampling events (site/day) per year")

ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per sampling event")

ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
  geom_point(size = 1, aes(alpha = obs)) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_alpha_continuous(name = "Observations") +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  theme_light(12) +
  labs(x = NULL,
       y = "Elevation (m x 10)") +
  ggtitle("Bumble bee observations through elevation and time")
```

# Bumble bee abundance
```{r message=FALSE, warning=FALSE}
bb_range <- visitation_sampling %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Range and peak abundance of each bumble bee sp.")

# Overall bumble bee abundance
bb_abund_sp <- visitation_sampling %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(visitation_sampling, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect))

bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species") +
  ggtitle("Relative abundances of bumble bee species")
```

# GAM diagnostics

## Bumble bee abundance GAM
```{r message=FALSE, warning=FALSE}
bb_abund_gam <- readRDS("../output/gam_max_bb_abund_01.rds")
summary(bb_abund_gam)
check.gamViz
```

## Floral abundance GAM
```{r message=FALSE, warning=FALSE}
fl_abund_gam <- readRDS("../output/gam_mean_fl_abund_01.rds")
summary(fl_abund_gam)
check.gamViz(fl_abund_gam)
```

>>>>

---
title: "sampling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale(locale = "en_US.UTF-8")
```

# Packages
```{r message=FALSE, warning=FALSE}
library(mgcViz)
library(lubridate)
library(patchwork)
library(tidyverse)
```

@@

## Visualize
```{r message=FALSE, warning=FALSE}
samplingplot1 <- ggplot(visitation_sampling, aes(x = year(date))) +
  geom_bar(fill = "gray60") +
  ylab("Total bumble bees\nobserved") +
  theme_light(11) +
  theme(axis.title.x = element_blank())

samplingplot2 <- ggplot(bb_obs_per_sample, aes(x = year)) +
  geom_bar(fill = "gray60") +
  ylab("Total samples") +
  theme_light(11) +
  theme(axis.title.x = element_blank())

samplingplot3 <- ggplot(bb_obs_per_sample, aes(x = year, y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Bumble bees\nobserved / sample") +
  #xlab(NULL) +
  theme_light(11) +
  theme(axis.title.x = element_blank())

samplingplot4 <- ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
  geom_point(size = 1, aes(alpha = obs)) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_alpha_continuous(name = "Bumble bees\nobserved") +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  theme_light(11) +
  labs(x = NULL,
       y = "Elevation (m x 10)")

samplingplot <- ((samplingplot1 / samplingplot2 / samplingplot3) | samplingplot4) +
  plot_layout(widths = c(1, 3)) + 
  plot_annotation(tag_levels = 'A')

ggsave("../ms_1/figures/samplingplot.pdf", height = 6, width = 10.5)
ggsave("../ms_1/figures/samplingplot.png", height = 6, width = 10.5)
```

### 2021-09-02 ms_1/Appendix_S1.Rmd
beginning ms1 revision

---
title: "Appendix S1"
output: pdf_document
---

@@

ggplot(bb_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Sampling events (site/day) per year")

ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per sampling event")

@@

## Bumble bee abundance GAM
```{r message=FALSE, warning=FALSE}
bb_abund_gam <- readRDS("../output/gam_max_bb_abund_01.rds")
summary(bb_abund_gam)
check.gamViz
```

>>>>

---
title: "The tree line ecotone punctuates the abundance and β-diversity of plants, pollinators, and their interactions"
subtitle: "Appendix S1: Analytical workflow"
author: 
  - Douglas B. Sponsler^[douglas.sponsler@uni-wuerzburg.de]
  - Katharina Kallnik 
  - Fabrice Requier 
  - Alice Classen 
  - A. Fabienne Maihoff
  - Johanna Sieger
  - Ingolf Steffan-Dewenter
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
#mainfont: Arial
---

@@

ggplot(bb_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Samplings (site*day) per year")

ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per sample")

@@

## Bumble bee abundance GAM
```{r message=FALSE, warning=FALSE}
bb_abund_gam <- readRDS("../output/gam_max_bb_abund_01.rds")
summary(bb_abund_gam)
check.gamViz(bb_abund_gam)
```
